---
title: "p8105_hw3_sp4170.rmd"
author: "Shihui Peng"
date: "2023-10-14"
output: github_document
---
```{r, echo=FALSE, message=FALSE}
library(tidyverse)
```


# Problem 2

Load the data from the p8105.datasets package.
```{r}
library(p8105.datasets)
data("brfss_smart2010")
```

## data cleaning

* First, do some data cleaning:
  * format the data to use appropriate variable names using `janitor::clean_names()`;
  * focus on the “Overall Health” topic using `filter()`
  * include only responses from “Excellent” to “Poor” also using `filter()`
  * organize responses as a factor taking levels ordered from “Poor” to “Excellent” using `mutate()` and `fct_relevel()`.
  
```{r}
brfss_df =
brfss_smart2010 |> 
  janitor::clean_names() |> 
  filter(
    topic == 'Overall Health', 
    response == 'Excellent' | response == 'Very good' | response =='Good' | response == 'Fair' | response == 'Poor') |> 
  mutate(
    response = fct_relevel(response, c('Poor', 'Fair', 'Good', 'Very good', 'Excellent'))
  )
```

i named the cleaned dataset as 'brfss_df'.

## answer questions

### Question 1

In 2002, which states were observed at 7 or more locations? What about in 2010?

```{r}
brfss_2002_df =
brfss_df |> 
  filter(year == 2002) |> 
  group_by(locationabbr) |> 
  summarise(n_locations = n_distinct(locationdesc)) |> 
  filter(n_locations >= 7)
```

* in 2002, states **`r pull(brfss_2002_df, locationabbr)`** were observed at 7 or more locations.

```{r}
brfss_2010_df =
brfss_df |> 
  filter(year == 2010) |> 
  group_by(locationabbr) |> 
  summarise(n_locations = n_distinct(locationdesc)) |> 
  filter(n_locations >= 7)
```

* in 2010, states **`r pull(brfss_2010_df, locationabbr)`** were observed at 7 or more locations.

### Question 2

requirement:

* Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. 
* Make a “spaghetti” plot of this average value over time within a state (that is, make a plot showing a line for each state across years – the geom_line geometry and group aesthetic will help).

```{r}
brfss_excellent_df =
brfss_df |> 
  filter(response == 'Excellent') |> 
  group_by(locationabbr) |> 
  mutate(
    mean_data_value = mean(data_value, na.rm = TRUE)
  ) |> 
  select(year, locationabbr, mean_data_value) |> 
  distinct()

brfss_excellent_df |> 
  ggplot(aes(x = year, y = mean_data_value)) +
  geom_line(aes(group = locationabbr, color = locationabbr)) +
  labs(
    title = "Average data value over time within states"
  )
```

* here, i first create a dataset `brfss_excellent_df` following the requirement. This new dataset only have 3 variables, which are year, locationabbr, and mean_data_value and is limited to Excellent responses. Then i do a spaghetti plot using geom_line() to show average data value over time within states.

### Question 3

Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.

```{r}
brfss_df |> 
  filter(
    locationabbr == 'NY',
    year == 2006 | year == 2010) |> 
  group_by(locationdesc) |> 
  ggplot(aes(x = data_value, fill = response)) +
  geom_density(alpha = 0.3) +
  facet_grid(. ~ year)+
  labs(
    title = "Distribution of data value for response among NY locations"
  )
```

