---
title: "p8105_hw3_sp4170.rmd"
author: "Shihui Peng"
date: "2023-10-14"
output: github_document
---
load libraries and packages that will be used. Then, set options using code provided by Jeff in the lecture.
```{r, echo=FALSE, message=FALSE}
library(tidyverse)
library(p8105.datasets)
data("brfss_smart2010")

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


# Problem 2

## data cleaning

* First, do some data cleaning:
  * format the data to use appropriate variable names using `janitor::clean_names()`;
  * focus on the “Overall Health” topic using `filter()`
  * include only responses from “Excellent” to “Poor” also using `filter()`
  * organize responses as a factor taking levels ordered from “Poor” to “Excellent” using `mutate()` and `fct_relevel()`.
  
```{r}
brfss_df =
brfss_smart2010 |> 
  janitor::clean_names() |> 
  filter(
    topic == 'Overall Health', 
    response == 'Excellent' | response == 'Very good' | response =='Good' | response == 'Fair' | response == 'Poor') |> 
  mutate(
    response = fct_relevel(response, c('Poor', 'Fair', 'Good', 'Very good', 'Excellent'))
  )
```

i named the cleaned dataset as 'brfss_df'.

## answer questions

### Question 1

In 2002, which states were observed at 7 or more locations? What about in 2010?

```{r}
brfss_2002_df =
brfss_df |> 
  filter(year == 2002) |> 
  group_by(locationabbr) |> 
  summarise(n_locations = n_distinct(locationdesc)) |> 
  filter(n_locations >= 7)
```

* in 2002, states **`r pull(brfss_2002_df, locationabbr)`** were observed at 7 or more locations.

```{r}
brfss_2010_df =
brfss_df |> 
  filter(year == 2010) |> 
  group_by(locationabbr) |> 
  summarise(n_locations = n_distinct(locationdesc)) |> 
  filter(n_locations >= 7)
```

* in 2010, states **`r pull(brfss_2010_df, locationabbr)`** were observed at 7 or more locations.

### Question 2

requirement:

* Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. 
* Make a “spaghetti” plot of this average value over time within a state (that is, make a plot showing a line for each state across years – the geom_line geometry and group aesthetic will help).

```{r}
brfss_excellent_df =
brfss_df |> 
  filter(response == 'Excellent') |> 
  group_by(locationabbr) |> 
  mutate(
    mean_data_value = mean(data_value, na.rm = TRUE)
  ) |> 
  select(year, locationabbr, mean_data_value) |> 
  distinct()

brfss_excellent_df |> 
  ggplot(aes(x = year, y = mean_data_value)) +
  geom_line(aes(group = locationabbr, color = locationabbr)) +
  labs(
    title = "Average data value over time within states"
  )
```

* here, i first create a dataset `brfss_excellent_df` following the requirement. This new dataset only have 3 variables, which are year, locationabbr, and mean_data_value and is limited to Excellent responses. Then i do a spaghetti plot using geom_line() to show average data value over time within states.

### Question 3

Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.

```{r}
brfss_df |> 
  filter(
    locationabbr == 'NY',
    year == 2006 | year == 2010) |> 
  group_by(locationdesc) |> 
  ggplot(aes(x = data_value, fill = response)) +
  geom_density(alpha = 0.3) +
  facet_grid(. ~ year)+
  labs(
    title = "Distribution of data value for response among NY locations"
  )
```

* here, i first filter the dataset and then do a density plot and use `facet_grid()` to make it as a 2-panel plot.


# Problem 3

## data loading, tidying, and merging and organizing

Load, tidy, merge, and otherwise organize the data sets. 
* include all originally observed variables
* exclude participants less than 21 years of age, and those with missing demographic data
* and encode data with reasonable variable classes (i.e. not numeric, and using factors with the ordering of tables and plots in mind).

```{r}
accel_df = 
  read_csv("data/nhanes_accel.csv") |> 
  janitor::clean_names() |> 
  pivot_longer(
    min1:min1440,
    names_to = 'time',
    values_to = 'mims'
  )

covar_df =
  read_csv("data/nhanes_covar.csv", skip = 4) |> 
  janitor::clean_names()

final_df =
  left_join(covar_df, accel_df, by = 'seqn') |> 
  drop_na(sex, age, bmi, education) |> 
  filter(age >= 21) |> 
  mutate(
    education = case_match(education, 1 ~ 'less_than_high_school', 2 ~ 'high_school_equivalent', 3 ~ 'more_than_high_school'),
    sex = case_match(sex, 1 ~ 'male', 2 ~ 'female'),
    education = fct_relevel(education, c('less_than_high_school', 'high_school_equivalent', 'more_than_high_school'))
  )
```

## Questions

### Question 1

* Produce a reader-friendly table for the number of men and women in each education category
* create a visualization of the age distributions for men and women in each education category. Comment on these items.

```{r}
final_df |> 
  select(sex, education, age) |> 
  group_by(sex, education) |> 
  summarise(n_obs = n()) |>
  knitr::kable()

final_df |> 
  select(sex, education, age) |> 
  ggplot(aes(x = age, color = sex))+
  geom_histogram(binwidth = 5, alpha = 0.3) +
  facet_grid(. ~ education) +
  labs(
    title = 'Age distributions by sex and education category'
  )
```

* 1st part code
  * here, i group the dataset by sex and education category using `group_by()` and then use `summarise()` to count how many observations in each single group (name this variable as n_obs). Then i use `knitr::kable()` to produce a reader-friendly table.
  * *comments*: 
    * In both sex groups, most participants are educated more than high school, while more females get education more than high school than males.
    * Among females, least participants are educated as high school equivalent.
    * Among males, least participants are educated as less than high school.
    * for group of getting less than high school education and getting more than high school education, counts of females are larger than of males.
  
* 2nd part code
  * *comments*: 
    * among those with less than high school education, most females aged around 70 and most males aged around 45 and 70.
    * among those with high school equivalent education, most females aged around 80 and most males aged around 55.
    * among those with more than high school education, most females aged around 30 and most males aged around 30.
    * we can see that the age distribution of females are more skewed for those who received more than high school education, while of males are more average across age.
    * for group of getting less than high school education and getting more than high school education, counts of females are larger than of males.
    
### Question 2

* Using your tidied dataset, aggregate across minutes to create a total activity variable for each participant. 
* Plot these total activities (y-axis) against age (x-axis); your plot should compare men to women and have separate panels for each education level. Include a trend line or a smooth to illustrate differences. Comment on your plot.

```{r}
final_df |>
  group_by(seqn) |> 
  mutate(sum_activity = sum(mims)) |> 
  select(seqn, sex, age, education, sum_activity) |> 
  distinct() |> 
  ggplot(aes(x = age, y = sum_activity, color = sex)) +
  geom_smooth() +
  facet_grid(. ~ education) +
  labs(
    title = 'Total activity over the day by sex and education'
  )
```

I include a smooth in my plot.

* *comments*:
  * for each education group, the sum_activity would decrease when ages increase for both genders.
  * for education group that is high school equivalent and more than high school, in most age, the sum_activity for females surpass males.
  * for education group that is less than high school, before around 40 years old, females have sum_activity greater than males. after around 40 years old, the situation reverses.
  * based on the plots, the group of less than high school education and the group of high school equivalent education have wider variations than the group of more than high school education.
  








